---
title: "Arrange the rows in descending/ascending order "
author: "Siddharth"
date: "`r Sys.Date()`"
output: html_document
---

# load the library

```{r}


library(nycflights13)
library(tidyverse)
library(gapminder)

```

# check few rows of data set

```{r}

glimpse(gapminder)

```

# Arrange the countries in ascending order of life expentancy

```{r}

gapminder %>% 
  arrange(lifeExp) %>%
  view()

```

# Arrange the countries in descending order of life expectancy

```{r}

gapminder %>% 
  arrange(desc(lifeExp))%>%
  view()

```

# Arrange the countries in ascending order of life expectancy and population.

```{r}

gapminder %>%
  arrange(lifeExp,pop) %>%
  view()

```

# Arrange the countries in descending order of life expectancy and population.

```{r}

gapminder %>%
  arrange(desc(lifeExp),desc(pop)) %>%
  view()

```

# Arrange the countries in ascending order of life expectancy and descending order of population.

```{r}
gapminder %>%
  arrange(lifeExp,desc(pop)) %>%
  view()

```

# load alcohol dataset

```{r}

setwd("E:\\Training Modules\\Dplyr Package Videos\\Introduction to Dplyr Package\\Session 5 - Arrange Rows")

drinks <- read.csv("drinks.csv",stringsAsFactors = FALSE)

#drinks <- read.csv("https://raw.githubusercontent.com/siddharth-sahasrabudhe/Dplyr-Package-Videos/main/drinks.csv")


```

# check few rows of data set

```{r}

glimpse(drinks)

```

# Arrange the countries in asscending order of beer servings

```{r}
drinks %>%
  arrange(beer_servings)%>%
  view()
```

# Arrange the countries in asscending order of beer serving and spirit servings

```{r}

drinks %>% 
  arrange(beer_servings,spirit_servings)%>%
  view()

```

# Arrange the countries in descending order of wine serving and spirit servings

```{r}
drinks %>% 
  arrange(desc(wine_servings),desc(spirit_servings))%>%
  view()

```

# Arrange in ascending order with one variable, and in descending order with another

```{r}

drinks %>% arrange(beer_servings,desc(wine_servings))%>%
  view()

```

# Order the rows with NA values

```{r}

flights_jan <- flights %>% 
  filter(month==1, day==1) 

sum(is.na(flights_jan))

flights_jan  %>% 
  filter(is.na(dep_delay)) %>%
  view()


flights_jan %>% arrange(dep_delay)%>%
  view()

  flights_jan %>% arrange(desc(arr_delay))%>%
  view()

```

## Summary:

-   `arrange()` function is used to sort out the rows in **ascending** order. The first argument to `arrange()` function is data-set followed by the name of the variable by which the rows are sorted out.

-   `arrange(desc())` method is used to sort out the rows in **descending** order.

-   If the first variable has multiple observations with the same value, the value of the second variable is used to arrange the rows.

-   Rows with `NA` values are always positioned at last.
