---
title: "Create new columns in data using mutate() function"
output: html_document
date: "2023-09-26"
---

## load the libraries

```{r}

library(tidyverse)
library(nycflights13)

```

## Check few rows of the data set

```{r}

head(flights,n=20)
?flights

```

## Create a sub-set of the data

```{r}

flights_subset <- flights %>%
  select(carrier,origin,dest,distance,ends_with("delay"),air_time) %>%
  view()

```

## create a new variable - total_travel_time = air_time + dep_delay + arr_delay

```{r}

flights_subset %>%
  mutate(total_travel_time = arr_delay + dep_delay + air_time) %>%
  view()

```

## Modify the existing variable - convert distance to km

```{r}

flights_subset %>%
  mutate(distance = distance*1.6) %>%
  view()

```

## create more than one variable using mutate()

```{r}
flights_subset %>%
  mutate(total_travel_time = arr_delay + dep_delay + air_time,
         total_delay = total_travel_time - air_time) %>%
  view()

```

## Display only the new columns

```{r}
flights_subset %>%
  transmute(total_travel_time = arr_delay + dep_delay + air_time,
         total_delay = total_travel_time - air_time) %>%
  view()

```

## Using mutate() on groups

```{r}

flights_subset %>%
  group_by(carrier)%>%
  mutate(total_travel_time = arr_delay + dep_delay + air_time,
         total_delay = total_travel_time - air_time,
         mean_delay = mean(total_delay,na.rm=TRUE)) %>%
  view()

```

## re-order the position of the new columns

```{r}

flights_subset %>%
  group_by(carrier)%>%
  mutate(total_travel_time = arr_delay + dep_delay + air_time,
         total_delay = total_travel_time - air_time,
         mean_delay = mean(total_delay,na.rm=TRUE),.after = carrier) %>%
    view()

```

## Hide un-wanted columns

```{r}
flights_subset %>%
  group_by(carrier)%>%
  mutate(distance=NULL,total_travel_time = arr_delay + dep_delay + air_time,
         total_delay = total_travel_time - air_time,
         mean_delay = mean(total_delay,na.rm=TRUE),.after = carrier) %>%
    view()

```

# Summary:

-   `mutate()` creates new columns that are functions of existing variables. It can also modify and delete existing columns (by setting their value to `NULL`).

-   The `mutate()` operation may yield different results on grouped data because the expressions are computed within groups.

-   By default, new columns are placed on the far right. You can change the order of columns using `.before` or `.after` argument
