---
title: "slice() function to slice the data "
author: "Siddharth"
date: "`r Sys.Date()`"
output: html_document
---

The `slice( )` function, as the name suggests, slices the data set into various cross sections. It selects, removes, duplicates rows by their location. The `slice()`function has a close resemblance with `arrange()` function, but has more options to sort the rows based on its integer location.

## load the library

```{r}

library(tidyverse)

```

## view the complete data set

```{r}

mtcars %>%
  view()

```

## Print out the last row of data set

```{r}
 mtcars %>%
  slice(n())


```

## Print out the first row of data set

```{r}

 mtcars %>%
  slice(1)

```

## Print out the first rows based on existing order

```{r}
 mtcars %>%
  slice_head(n=5)

```

## Print out the last rows based on existing order

```{r}
 mtcars %>%
  slice_tail(n=5)

```

## Which are the top 10 car models as per miles per gallon?

```{r}

mtcars %>%
  slice_max(mpg,n=10)

```

## Which are the top 5 car models as per the number of engine cylinders?

```{r}
mtcars %>%
  slice_max(cyl,n=5)

```

## Which are the 10 car models with lower weight?

```{r}

mtcars %>%
  slice_min(wt,n=10)

```

# Randomly select the observations from the data set.

```{r}

mtcars %>%
  slice_sample(n=10,replace = TRUE)

```

## Drop first 5 observations from the data set.

```{r}

mtcars %>% slice(-c(1:5))
```

## Drop last 5 observations from the data set.

```{r}

mtcars %>%
  slice(c(n()-5:n()))

```

## Select observations based on weight measure

```{r}

mtcars %>%
  slice_sample(weight_by = disp,n=5)

```

## Select observations among the grouped variables

```{r}
mtcars %>%
  group_by(cyl)%>%
    slice_head(n=2)%>%


```

## Summary:

-   `slice()` lets you index rows by their (integer) locations.

-   `slice_head()` and `slice_tail()` select the first or last rows.

-   `slice_sample()` randomly selects rows.

-   `slice_min()` and `slice_max()` select rows with the smallest or largest values of a variable.
