---
title: "Aggregate the data using group_by() and summrise() function"
author: "Siddharth"
date: "`r Sys.Date()`"
output: html_document
---

`dplyr` package provides the `group_by()` function to
group data based on specific variables and the `summarize()` function to calculate summary statistics for each group.

## load the library

```{r}
library(tidyverse)

```

## load the IMDB data.

```{r}

Movies_data <- read.csv("IMDB Movies.csv")

# Movies_data <- read.csv("https://raw.githubusercontent.com/siddharth-sahasrabudhe/Dplyr-Package-Videos/main/IMDB%20Movies.csv")

```

## Check few rows of the dataset

```{r}

 glimpse(Movies_data)


```

## Generate the summary statistics of Gross Revenue

```{r}

Movies_data %>% summarise(Mean_revenue = sd(Gross.Revenue),n=n())


```

## Generate the summary statistics of Gross Revenue by Genre and display top 10 Genre by Mean Revenue

```{r}

Movies_data %>% 
  group_by(Genre) %>%
  summarise(Mean_revenue = mean(Gross.Revenue),variance = sd(Gross.Revenue),n=n()) %>%
  slice_max(Mean_revenue,n=10)
  

```

## Generate the summary statistics of Gross Revenue and profit by Gennre and Country 

```{r}

Movies_data %>%
  group_by(Color.B.W,Genre)%>%
  summarise(Mean_revenue = mean(Gross.Revenue),profit = (`Gross.Revenue`-Budget) ,n=n())%>%
  filter(profit < 0)

```
 
## Ungroup the data back into the original one

```{r}

Movies_data %>%
    ungroup()

```

## Summary:

-   `group_by()` converts the data set into a grouped data where operations are performed "by group". `ungroup()`        removes grouping.

-   `summarise()` creates a new data frame. It returns one row for each combination of grouping variables.

